---
title: "DATA605_Assignment3"
author: "Harpreet Shoker"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### (1) What is the rank of the matrix A?
$$ A = \begin{bmatrix}1 & 2 & 3 & 4 \\ -1 & 0 & 1 & 3 \\0 & 1 & -2 & 1 \\5 & 4 & -2 & -3\end{bmatrix}$$ 
```{r }
library(pracma)
A <- matrix(c(1,2,3,4,-1,0,1,3,0,1,-2,1,5,4,-2,-3),nrow = 4,byrow = TRUE)
A
rref(A)
Rank(A)
```
We see 4 pvot rows when reduced to row echolen form using rref()
checking with Rank() function

Both gives result 4

### (2) Given an mxn matrix where m > n, what can be the maximum rank? The mini- mum rank, assuming that the matrix is non-zero?
Answer :- Maximum rank = m rows 
Minimum rank >= 1

### (3) What is the rank of matrix B?
$$ A = \begin{bmatrix}1 & 2 & 1 \\ 3 & 6 & 3 \\ 2 & 4 & 2\end{bmatrix}$$
```{r }
B <- matrix(c(1,2,1,3,6,3,2,4,2),nrow = 3,byrow = TRUE)
B
rref(B)
Rank(B)
```

Ranks is 1

### (4) Compute the eigenvalues and eigenvectors of the matrix A. You'll need to write out the characteristic polynomial and show your solution.

$$ A = \begin{bmatrix}1 & 2 & 3 \\0 & 4 & 5 \\0 & 0 & 6 \end{bmatrix}$$ 
$$ \lambda\,I_3 = \begin{bmatrix}\lambda & 0 & 0 \\0 & \lambda & 0 \\0 & 0 & \lambda \end{bmatrix}$$ 
$$ det(A-\lambda\,I_n)=0$$ 
$$ det\,\begin{bmatrix}1-\lambda & 2 & 3 \\0 & 4-\lambda & 5 \\0 & 0 & 6-\lambda \end{bmatrix} = 0$$ $$(1-\lambda)(4-\lambda)(6-\lambda)=0$$ $$ Eigenvalues\,of\,A:$$ $$\lambda=1,\, \lambda=4,\, \lambda=6$$

Calculating Eigenvectors using R

For λ1=1 
```{r }
A <- matrix(c(1,2,3,0,4,5,0,0,6), nrow=3, byrow=TRUE)
I <- diag(3)
Eigen1 <- A - (1*I)
rref(Eigen1)
```

So from the above reduced form we see
x3 = x2 = 0 and x1 is free
So eigenspace = $$E_{=1}=\begin{bmatrix}1 \\ 0 \\ 0\end{bmatrix}$$

For λ2=4
```{r }
A <- matrix(c(1,2,3,0,4,5,0,0,6), nrow=3, byrow=TRUE)
I <- diag(3)
Eigen2 <- A - (4*I)
rref(Eigen2)
```

Here we see x3 = 0 , x1 - 0.66(x2)= 0
so x2 = 1.5 (x1)
Hence eigenspace = $$E_{=4}=\begin{bmatrix}1 \\ 1.5 \\ 0\end{bmatrix}$$


For λ=6
```{r }
A <- matrix(c(1,2,3,0,4,5,0,0,6), nrow=3, byrow=TRUE)
I <- diag(3)
Eigen3 <- A - (6*I)
rref(Eigen3)
```

 
x2 - (2.5)x3 = 0 , x2 = 2.5(x3)
X1 - (1.6)x3 = 0 , x1 = (1.6)x3

Hence eigenspace = $$E_{=4}=\begin{bmatrix}1.6 \\ 2.5 \\ 1\end{bmatrix}$$





```